---
title: "Seasonality in NOCs"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
runtime: shiny
resource_files:
- data/RTRA1446483_stat1115NOCp2.csv
- data/RTRA3040048_stat2125NOCp1.csv
- data/RTRA4340555_stat2125NOCp2.csv
- data/RTRA5495420_stat1620NOCp2.csv
- data/RTRA6895467_stat1115NOCp1.csv
- data/RTRA8270443_stat1620NOCp1.csv
- mapping/Occupational Characteristics w skills interests wages.xlsx
---
    
```{r setup, include=FALSE}
library(shiny)
library(plotly)
library(htmlwidgets)
library(tidyverse)
library(here)
library(fpp3)
library(janitor)
library(conflicted)
library(readxl)
conflicts_prefer(dplyr::filter)
my_dt <- function(tbbl) {
  DT::datatable(tbbl,
                filter = 'top',
                extensions = "Buttons",
                rownames = FALSE,
                options = list(
                  columnDefs = list(list(className = "dt-center", targets = "_all")),
                  paging = TRUE,
                  scrollX = TRUE,
                  scrollY = TRUE,
                  searching = TRUE,
                  ordering = TRUE,
                  dom = "Btip",
                  buttons = list(
                    list(extend = "csv", filename = "occupation_features"),
                    list(extend = "excel", filename = "occupation_features")
                  ),
                  pageLength = 20
                )
  )
}
noc_names <- read_excel(here("mapping", 
                             "Occupational Characteristics w skills interests wages.xlsx"))|>
  clean_names()|>
  select(noc, description)|>
  mutate(NOC_5=str_remove_all(noc, "#"))

tbbl <- vroom::vroom(here("data",list.files(here("data"))))|>
  filter(!is.na(NOC_5),
         !is.na(SYEAR),
         !is.na(SMTH))|>
  mutate(date=yearmonth(ymd(paste(SYEAR,SMTH,"01", sep = "/"))),
         NOC_5=as.character(NOC_5))|>
  select(-SYEAR,-SMTH)|>
  filter(date<yearmonth(today())-month(1))|>
  left_join(noc_names)|>
  select(-noc)|>
  unite(NOC_5, NOC_5, description, sep=": ")|>
  tsibble(key=NOC_5, index = date)

feats <- tbbl|>
  features(`_COUNT_`, feat_stl)
```

Trend vs. Seasonality 
=====================================  

Column {data-width=300}
-----------------------------------------------------------------------

### Click on a point to investigate

```{r, fig.retina=2}
js <- "
function(el, x) {
  el.on('plotly_click', function(d) {
    var point = d.points[0];
    Shiny.setInputValue('point', {text: point.text});
  });
}"
renderPlotly({
  plt <- ggplot(feats, aes(trend_strength,
                        seasonal_strength_year,
                        text=NOC_5))+
  geom_point(alpha=.25)+
  scale_y_continuous(trans="log10")
  
  ggplotly(plt, tooltip = "text")|>
    onRender(js)
})
```


Column {data-width=500}
-----------------------------------------------------------------------

### STL decompostition of `r renderUI({input$point$text})`

```{r, fig.retina=2}
renderPlot({
  req(input$point)
tbbl |>
  filter(NOC_5==input$point$text)|>
  model(stl = STL(`_COUNT_`, t.window=13, robust=TRUE))|>
  components()|>
  autoplot()+ 
  theme(text=element_text(size=15))
})
```


Column {data-width=500}
-----------------------------------------------------------------------

### Subseries Plot of `r renderUI({input$point$text})`

```{r, fig.retina=2}
renderPlot({
  req(input$point)
tbbl |>
  filter(NOC_5==input$point$text)|>
  gg_subseries(`_COUNT_`)+
  theme(text=element_text(size=15))
})
```



### Season Plot of `r renderUI({input$point$text})`

```{r, fig.retina=2}
renderPlot({
  req(input$point)
tbbl |>
  filter(NOC_5==input$point$text)|>
  gg_season(`_COUNT_`, labels = "both")+
  theme(text=element_text(size=15))
})
```

Description of features
======================================

Quoting from Forecasting: Principles and Practice (3rd ed) by Rob J Hyndman and George Athanasopoulos:

-   A time series decomposition can be used to measure the strength of trend and seasonality in a time series. 
-   Recall that the decomposition is written as $y_t=T_t+S_t+R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component. 
-   For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
-   Therefore $Var(R_t)/Var(T_t+R_t)$ should be relatively small. But for data with little or no trend, the two variances should be approximately the same. 
-   So we define the strength of trend as: $$FT=\max\left(0,1−\frac{Var(R_t)}{Var(T_t+R_t)}\right)$$. 
-   This will give a measure of the strength of the trend between 0 and 1. Because the variance of the remainder might occasionally be even larger than the variance of the seasonally adjusted data, we set the minimal possible value of $FT$ equal to zero.

-   The strength of seasonality is defined similarly, but with respect to the detrended data rather than the seasonally adjusted data: $$FS=\max\left(0,1−\frac{Var(R_t)}{Var(S_t+R_t)}\right)$$.
-   A series with seasonal strength $FS$ close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $FS$ close to 1 because $Var(R_t)$ will be much smaller than $Var(S_t+R_t)$.


The feature data:
=====================================  

```{r}
my_dt(feats[,1:3])
```




